---
title: "Aides Drug"
author:  Aylin Torres
date: 10/08/24
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

$h_0$: The vaccine is not effective in preventing women from contracting aids.
$h_A$: The vaccine is effective in preventing women from contracting aids.


```{r}
women <- triUmp[triUmp$gender=='F',]
women 
```


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=women)
barchartGC(~infected + treatment,data=women, type="percent")
```

among women who contracted aids, most of them had undergone treatment. Therefore the treatment does not appear to be effective.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <-xtabs(~infected + treatment, data=women)
rowPerc(table1)
colPerc(table1)
```
Among those who got aids,58.82% of them received treatment.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fishers test shows that p-value is 0.00006596,which is lower than the acceptable p-value of 0.5 so we will rejected the null hypothesis. It also states that women who gets the vaccine are 1.543902 times more likely to get aids than who did not get vaccine.Therefore the treatment is not effective in preventing aids in women.

# Discussion and Conclusion


Ho The vaccine is not effective in preventing females from contracting aids.
H, The vaccine is effective in preventing females from contracting aids.
We will reject the null hypothesis because.
According to the barchart, among the women who have contracted alds, most received the vaccine. This chart shows the vaccine is not effective in preventing the disease.
The numerical results shows that the women who have received treatment about 58.82 percent of women received the vaccine but still contracted aids. While those who received the placebo only 41.18 percent contracted aids.
The fishers test shows that the p-value is 0.00006596, which is lower than the acceptable p value of 0.5 so we will reject the null hypothesis.It also states that women who get the vaccine are 1.543902 times more likely to get aids than those who did not get the vaccine.
Therefore the treatment is not effective in preventing alds in women.




<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
